Ensemble labels

### 2021-03-06 docs/ensemble_labels.Rmd
Some updates and applying OSM to clean labels

---
title: "Ensemble labels"
author: "Lei Song"
date: "`r Sys.Date()`"
output: 
  rmdformats::readthedown:
    highlight: pygments
    toc_depth: 5
---

@@

#### Select out sub-grids

@@

##### [ARCHIVED] Get large amount of weak labels.
Based on the prior experiments, this method didn't do well. The train dataset is "under-representative" in terms of DL training (Validate loss curve is below the train loss significantly). This might be caused by either the super imbalance in both train and validate dataset, or the noises in the labels.

@@

#### Split training, validation, and test

@@

```{r labels_img}
library(glue)
tiles <- st_read('data/intermid/tiles_grids.geojson',
                 quiet = T)
lc_labels <- rast('data/intermid/lc_labels_exd_val.tif')

plt_path <- '/Volumes/elephant/plt_nicfi'
plt_nms <- list.files(plt_path, full.names = T)
labels_path <- '/Volumes/elephant/labels'

mclapply(unique(catalog_train_val_test$tile), 
         function(tile_nm){
  tiles_each <- catalog_train_val_test %>% 
    filter(tile == tile_nm)
  lc_labels_each <- crop(lc_labels, tiles %>% 
                           filter(tile == tile_nm) %>% 
                           st_union() %>% 
                           st_buffer(0.01) %>% 
                           st_sf() %>% vect())
  nm <- grep(tile_nm, plt_nms, value = TRUE)[1]
  temp <- rast(nm)[[1]] %>% 
    aggregate(fact = 4096/width_chips)
  vals <- matrix(1:(width_chips^2), 
                 nrow = width_chips, 
                 ncol = width_chips) %>% t()
  vals <- vals[nrow(vals):1, ]
  values(temp) <- vals
  lapply(1:nrow(tiles_each), function(n){
    tile_each <- tiles_each %>% slice(n)
    temp_each <- temp
    index <- tile_each$index
    temp_each[temp_each != index] <- NA
    temp_each <- terra::trim(temp_each) %>% 
      disaggregate(fact = 4096/width_chips)
    label_each <- resample(lc_labels_each, temp_each, method = 'near')
    label_each[label_each == 8] <- 7
    label_each[label_each == 9] <- 8
    writeRaster(label_each, 
                file.path(labels_path, 
                          glue('labels_{tile_nm}_{index}.tif')),
                wopt = list(datatype = 'INT1U',
                        gdal=c("COMPRESS=LZW")))
    rm(tile_each, temp_each, label_each, index);gc()
  })
  rm(tiles_each, lc_labels_each, nm, temp, vals);gc()
}, mc.cores = 12)
```

>>>>

---
title: "Ensemble labels"
author: "Lei Song"
date: "`r Sys.Date()`"
output: 
  rmdformats::readthedown:
    highlight: pygments
    toc_depth: 6
---

@@

#### Select out sub-grids
##### Statistic

@@

##### Reselect
```{r reselect}
grids_pure_count$n[4:nrow(grids_pure_count)] <- ceiling(mean(grids_pure_count$n[1:3]))
tiles_grids_selected <- do.call(rbind, 
                                lapply(grids_pure_count$name, 
                                       function(nm){
    message(nm)
    n_grids <- grids_pure_count %>% 
      filter(name == nm) %>% 
      pull(n)
    set.seed(123)
    tile_select <- tiles_labels %>% 
      dplyr::select(tile, index, all_of(nm)) %>% 
      filter(!is.na(!!rlang::sym(nm))) %>% 
      filter((!!rlang::sym(nm)) >= 0.8) %>% 
      sample_n(min(n_grids, nrow(.)))
    data.frame(name = nm, 
               tile = tile_select$tile, 
               index = tile_select$index)
}))
```

#### Split training and validation

```{r split_pure}
set.seed(1000)
tiles_grids_train <- tiles_grids_selected %>% 
  group_by(name) %>% sample_frac(0.7) %>% 
  ungroup() %>% dplyr::select(-name) %>% 
  unique() %>% mutate(usage = 'train') %>%
  dplyr::select(tile, index, usage)
tiles_grids_valid <- tiles_grids_selected %>% 
  anti_join(tiles_grids_train, 
            by = c('tile', 'index')) %>% 
  group_by(name) %>% ungroup() %>% 
  dplyr::select(-name) %>% 
  unique() %>% mutate(usage = 'validate') %>%  
  dplyr::select(tile, index, usage)
catalog_train_val <- rbind(tiles_grids_train, 
                               tiles_grids_valid)
write.csv(catalog_train_val, 
          'data/intermid/catalog_train_val_pure.csv',
          row.names = F)
```

#### [ARCHIVED] Get large amount of weak labels.
Based on the prior experiments, this method didn't do well. The train dataset is "under-representative" in terms of DL training (Validate loss curve is below the train loss significantly). This might be caused by either the super imbalance in both train and validate dataset, or the noises in the labels.

@@

#### Split training, validation, and test
Select test dataset for final round.

@@

```{r labels_img}
library(glue)

# Make the script consistent
catalog_train_val_test <- catalog_train_val
tiles <- st_read('data/intermid/tiles_grids.geojson',
                 quiet = T)
lc_labels <- rast('data/intermid/lc_labels_exd_val.tif')

plt_path <- '/Volumes/elephant/plt_nicfi'
plt_nms <- list.files(plt_path, full.names = T)
labels_path <- '/Volumes/elephant/labels_pure'

lapply(unique(catalog_train_val_test$tile), 
         function(tile_nm){
  tiles_each <- catalog_train_val_test %>% 
    filter(tile == tile_nm)
  lc_labels_each <- crop(lc_labels, tiles %>% 
                           filter(tile == tile_nm) %>% 
                           st_union() %>% 
                           st_buffer(0.01) %>% 
                           st_sf() %>% vect())
  nm <- grep(tile_nm, plt_nms, value = TRUE)[1]
  temp <- rast(nm)[[1]] %>% 
    aggregate(fact = 4096/width_chips)
  vals <- matrix(1:(width_chips^2), 
                 nrow = width_chips, 
                 ncol = width_chips) %>% t()
  vals <- vals[nrow(vals):1, ]
  values(temp) <- vals
  lapply(1:nrow(tiles_each), function(n){
    tile_each <- tiles_each %>% slice(n)
    temp_each <- temp
    index <- tile_each$index
    temp_each[temp_each != index] <- NA
    temp_each <- terra::trim(temp_each) %>% 
      disaggregate(fact = 4096/width_chips)
    label_each <- resample(lc_labels_each, temp_each, method = 'near')
    label_each[label_each == 8] <- 7
    label_each[label_each == 9] <- 8
    writeRaster(label_each, 
                file.path(labels_path, 
                          glue('labels_{tile_nm}_{index}.tif')),
                wopt = list(datatype = 'INT1U',
                        gdal=c("COMPRESS=LZW")))
    rm(tile_each, temp_each, label_each, index);gc()
  })
  rm(tiles_each, lc_labels_each, nm, temp, vals);gc()
})
```

### 2021-03-06 docs/ensemble_labels_machine.Rmd
Some updates and applying OSM to clean labels

### TANSIS mask
```{r}
url_cp <- 'https://files.osf.io/v1/resources/yrj3h/providers/osfstorage/5c28365528af78001a63ee6d?action=download&direct&version=1'
dir.create('data/tansis')
download.file(url_cp, 'data/tansis/cp_pred.zip')
unzip('data/tansis/cp_pred.zip', exdir = 'data/tansis')
```

@@

# Get the existing types
# Because not all types exist in the current study area
ids <- unique(c(unique(esaglc_new), 
                unique(fromglc_new), 
                1)) # For GFSAD and TANSIS
lc_types <- lc_types %>% filter(id %in% ids)
# Even though there is some pixels classified as snow/ice
# this class is too tiny in Tanzania,
# so we delete this class.
lc_types <- lc_types %>% filter(id != 10)

tansis_mask <- rast(list.files('data/tansis', full.names = T, pattern = '.tif$'))
tansis_mask <- tansis_mask$TZ__7
tansis_mask[is.na(tansis_mask)] <- 0
tansis_mask <- resample(tansis_mask != 0,
                        fromglc_new, 
                        method = 'near')

gfsad_mask <- resample(gfsad_new != 0,
                       fromglc_new, 
                       method = 'near')
crop_mask <- (tansis_mask + gfsad_mask) == 0
lapply(1:nrow(lc_types), function(n){
  lc <- lc_types %>% slice(n)
  message(lc$name)
  if (lc$id == 1){
    message('Just use GFSAD for cropland.')
    lc_lb_emb <- resample(gfsad_new == lc$id, 
                          fromglc_new, 
                          method = 'near')
    gc()
  } else if (lc$id == 6){
    message('Use all three products for water.')
    lc_lb_fromglc <- fromglc_new == lc$id
    lc_lb_esaglc <- resample(esaglc_new == lc$id, 
                             lc_lb_fromglc, 
                             method = 'near')
    lc_lb_gfsad <- resample(gfsad_new == lc$id, 
                            lc_lb_fromglc, 
                            method = 'near')
    lc_lb_emb <- (lc_lb_esaglc + 
                    lc_lb_fromglc + 
                    lc_lb_gfsad) >= 3
    lc_lb_emb <- lc_lb_emb * crop_mask
    rm(lc_lb_esaglc, lc_lb_fromglc, lc_lb_gfsad); gc()
  } else {
    message('Not use GFSAD for other types.')
    lc_lb_fromglc <- fromglc_new == lc$id
    lc_lb_esaglc <- resample(esaglc_new == lc$id, 
                             lc_lb_fromglc,
                             method = 'near')
    lc_lb_emb <- (lc_lb_esaglc + 
                    lc_lb_fromglc) == 2
    message('Skip the cropland area defined by gfsad and tansis.')
    lc_lb_emb <- lc_lb_emb * crop_mask
    # lc_lb_emb <- lc_lb_emb * gfsad_mask
    rm(lc_lb_esaglc, lc_lb_fromglc); gc()
  }

@@

For Tanzania case, there is very tiny Tundra and Snow/Ice type if any because Tanzania is a tropical country. The only possible is on the Q mountain.

@@

# save out
dst_path <- 'data/intermid'
fn <- file.path(dst_path, 'lc_labels.tif')
writeRaster(labels, fn, overwrite = TRUE,
              wopt = list(datatype = 'INT1U',
                          gdal=c("COMPRESS=LZW")))
fn <- file.path(dst_path, 'labels_extent.tif')
writeRaster(labels_ext, fn, overwrite = TRUE,
              wopt = list(datatype = 'INT1U',
                          gdal=c("COMPRESS=LZW")))
```

#### Select out sub-grids

Start-up zone: zone 5

```{r g1}
library(ggplot2)
library(tidyr)
library(parallel)
# Define the width of chips to cut
# In this case, 256 * 256 chip.
# The chip size is the trade-off between computing cost, 
# purity of ensemble labels, and the keep of geographic context.
width_chips <- 16
lc_labels <- rast('data/intermid/lc_labels.tif')
tiles <- st_read('data/geoms/tiles_nicfi.geojson',
                 quiet = T) %>% dplyr::select(tile)
ecozones <- st_read('data/intermid/clusters_final.geojson',
                    quiet = T)

tiles <- do.call(rbind, mclapply(1:nrow(tiles), function(n){
  tile <- tiles %>% slice(n)
  tiles_grids <- st_make_grid(tile, n = c(width_chips, width_chips)) %>% 
    st_sf() %>% arrange() %>% 
    mutate(tile = tile$tile,
           index = c(1:(width_chips^2)))
}, mc.cores = 12))
st_write(tiles, 'data/intermid/tiles_grids_ecozones_256.geojson')
tiles <- tiles %>% vect()

@@

tiles <- st_read('data/intermid/tiles_grids_ecozones_256.geojson')
tiles <- tiles %>% dplyr::select(-ecozone) %>% unique()
st_write(tiles, 'data/intermid/tiles_grids_256.geojson')

ecozones <- st_read('data/intermid/clusters_final.geojson',
                    quiet = T)
ecozone_5 <- ecozones %>% filter(region == 5)
tiles_zone5 <- tiles %>% 
    filter(st_intersects(ecozone_5, tiles, sparse = F)) %>% 
    vect()
st_write(tiles_zone5, 'data/intermid/tiles_grids_zone5.geojson')

# lc_types <- lc_types %>% filter(!id %in% c(5, 7, 9, 10))
tiles_labels_zone5 <- mclapply(1:nrow(tiles_zone5), function(n){
    labels_each <- crop(lc_labels, tiles_zone5[n, ])
    lb <- freq(labels_each) %>% data.frame() %>% 
        filter(value != 0) %>%
        mutate(count = count / (dim(labels_each)[1] * dim(labels_each)[2])) %>%
        left_join(., lc_types, by = c('value' = 'id')) %>%
        dplyr::select(-c(value, layer)) %>%
        pivot_wider(names_from = name, values_from = count) %>%
        mutate(tile = tiles_zone5[n, ]$tile,
               index = tiles_zone5[n, ]$index)
    if (nrow(lb) > 0) lb
}, mc.cores = 8) %>% bind_rows()

tiles_labels_zone5 <- tiles_labels_zone5 %>%
    dplyr::select(-c('Wetland', 'Bareland', 'Tundra'))

class_nm <- c('Cropland', 'Forest', 'Grassland', 
              'Shrubland', 'Water')
grids_pure_zone5 <- do.call(rbind, 
                      lapply(class_nm, 
                             function(nm){
                                 message(nm)
                                tile_select <- tiles_labels_zone5 %>% 
                                         dplyr::select(tile, index, all_of(nm)) %>% 
                                         filter(!is.na(!!rlang::sym(nm))) %>% 
                                         filter((!!rlang::sym(nm)) >= 0.9)
                                data.frame(name = nm, 
                                            tile = tile_select$tile, 
                                            index = tile_select$index,
                                            purity = tile_select %>% 
                                                dplyr::select((!!rlang::sym(nm))) %>% 
                                                pull)
                             }))

grids_pure_count <- grids_pure_zone5 %>% 
    group_by(name) %>% summarize(n = n()) %>% 
    arrange(n)

# This area covers most of the LC types, only miss enough built-up.
# Fortunately, built-up has very obvious and relatively same feature across the whole regions.
# so we could get some from other regions.
set.seed(123)
grids_labels_zone5 <- grids_pure_zone5 %>% 
    group_by(name) %>% 
    sample_n(250) %>% 
    ungroup()

# built-up
tile_select <- tiles_labels %>% 
    dplyr::select(tile, index, all_of('Urban/Built up')) %>% 
    filter(!is.na(!!rlang::sym(nm))) %>% 
    filter((!!rlang::sym(nm)) >= 0.5)
urbans <- data.frame(name = nm, 
           tile = tile_select$tile, 
           index = tile_select$index,
           purity = tile_select %>% 
               dplyr::select((!!rlang::sym(nm))) %>% 
               pull)
grids_labels_zone5 <- rbind(grids_labels_zone5, urbans)
grids_labels_zone5_json <- st_read('data/intermid/tiles_grids_256.geojson') %>% 
    right_join(grids_labels_zone5, by = c('tile', 'index'))
st_write(grids_labels_zone5_json, 'data/intermid/grids_labels_zone5.geojson')
```

@@

```{r split_pure}
set.seed(100)
grids_train <- grids_labels_zone5 %>% 
  group_by(name) %>% sample_frac(0.71) %>% 
  ungroup() %>% dplyr::select(-name) %>% 
  unique() %>% mutate(usage = 'train') %>%
  dplyr::select(tile, index, usage)

grids_valid <- grids_labels_zone5 %>% 
  anti_join(grids_train, 
            by = c('tile', 'index')) %>% 
  group_by(name) %>% ungroup() %>% 
  dplyr::select(-name) %>% 
  unique() %>% mutate(usage = 'validate') %>%  
  dplyr::select(tile, index, usage)

# In order to set the batch size easily for DL,
# We re-arrange the sites among train and validate
# Train 1020, so 1020 / 30 could be without remainder.
# Validate 410, so 410 / 10 could be without remainder.
set.seed(678)
ids <- sample(1:nrow(grids_valid), 
              size = (nrow(grids_valid) - (1020 - nrow(grids_train))) %% 10)
grids_valid <- grids_valid %>% slice(-ids)

set.seed(568)
ids <- sample(1:nrow(grids_valid), size = 1020 - nrow(grids_train))
grids_valid$usage[ids] <- 'train'

@@

write.csv(catalog_train_val, 
          'data/intermid/catalog_guess_zone5_train_val.csv',
          row.names = F)
```

@@

```
id          name
1       Cropland
2       Forest
3       Grassland
4       Shrubland
5       Urban/Built up (8)
6       Water
```

@@

# Make the script consistent
tiles <- st_read('data/intermid/tiles_grids_256.geojson',
                 quiet = T)
lc_labels <- rast('data/intermid/lc_labels.tif')

plt_path <- '/Volumes/elephant/plt_nicfi'
plt_nms <- list.files(plt_path, full.names = T)
labels_path <- '/Volumes/elephant/labels_guess_zone5'

mclapply(unique(catalog_train_val$tile),
         function(tile_nm){
  tiles_each <- catalog_train_val %>% 
    filter(tile == tile_nm)
  lc_labels_each <- crop(lc_labels, tiles %>% 
                           filter(tile == tile_nm) %>% 
                           st_union() %>% 
                           st_buffer(0.01) %>% 
                           st_sf() %>% vect())
  nm <- grep(tile_nm, plt_nms, value = TRUE)[1]
  temp <- rast(nm)[[1]] %>% 
    aggregate(fact = 4096/width_chips)
  vals <- matrix(1:(width_chips^2), 
                 nrow = width_chips, 
                 ncol = width_chips) %>% t()
  vals <- vals[nrow(vals):1, ]
  values(temp) <- vals
  lapply(1:nrow(tiles_each), function(n){
    tile_each <- tiles_each %>% slice(n)
    temp_each <- temp
    index <- tile_each$index
    temp_each[temp_each != index] <- NA
    temp_each <- terra::trim(temp_each) %>% 
      disaggregate(fact = 4096/width_chips)
    label_each <- resample(lc_labels_each, temp_each, method = 'near')
    # Remove wetland, bareland, tundra, snow/ice
    label_each[label_each %in% c(5, 7, 9, 10)] <- 0
    # re-define urban
    label_each[label_each == 8] <- 5

@@

plts <- merge(plt_quads_rain_season, plt_quads_dry_season, by = 'tile')
rm(plt_quads_rain_season, plt_quads_dry_season)
## Labels
lc_labels <- list.files('/Volumes/elephant/labels_guess_zone5',
                        full.names = T, pattern = '.tif$') %>% 
    data.frame(dir_label = .) %>% 
    mutate(name = str_extract(dir_label, 
                              '[0-9]+-[0-9]+_[0-9]+')) %>% 
    tidyr::separate(name, c('tile', 'index'), '_')
imgs <- merge(lc_labels, plts, by = 'tile')
# rm(plts, lc_labels)

catalog_train_val <- read.csv('data/intermid/catalog_guess_zone5_train_val.csv',
                                  stringsAsFactors = F) %>% 
    merge(imgs, by = c('tile', 'index')) %>% 
    mutate(name = paste(tile, index, sep = '-')) %>% 
    dplyr::select(name, usage, dir_gs, dir_os, dir_label)
write.csv(catalog_train_val, 'data/intermid/catalog_zone5_DL_guess_train_val.csv',
          row.names = F)
```

@@

```{r}
pred_catalog <- catalog_train_val %>% 
  mutate(tile = str_extract(str_extract(dir_gs, 
                                        '[0-9]+-[0-9]+.tif'), 
                            '[0-9]+-[0-9]+')) %>% 
  dplyr::select(tile, dir_gs, dir_os) %>% unique() %>% 
  filter(!tile %in% (urbans %>% dplyr::select(tile) %>% unique() %>% pull(tile)))
write.csv(pred_catalog, 'data/intermid/catalog_zone5_DL_guess_predict.csv',
          row.names = F)
```

>>>>

@@

# Get the existing types
# Because not all types exist in the current study area
ids <- unique(c(unique(esaglc_new), 
                unique(fromglc_new), 
                1)) # For GFSAD
lc_types <- lc_types %>% filter(id %in% ids)
# Even though there is some pixels classified as snow/ice
# this class is too tiny in Tanzania,
# so we delete this class.
lc_types <- lc_types %>% filter(id != 10)

gfsad_mask <- resample(gfsad_new == 0,
                       fromglc_new, 
                       method = 'near')
lapply(1:nrow(lc_types), function(n){
  lc <- lc_types %>% slice(n)
  message(lc$name)
  if (lc$id == 1){
    message('Just use GFSAD for cropland.')
    lc_lb_emb <- resample(gfsad_new == lc$id, 
                          fromglc_new, 
                          method = 'near')
    gc()
  } else if (lc$id == 6){
    message('Use all three products for water.')
    lc_lb_fromglc <- fromglc_new == lc$id
    lc_lb_esaglc <- resample(esaglc_new == lc$id, 
                             lc_lb_fromglc, 
                             method = 'near')
    lc_lb_gfsad <- resample(gfsad_new == lc$id, 
                            lc_lb_fromglc, 
                            method = 'near')
    lc_lb_emb <- (lc_lb_esaglc + 
                    lc_lb_fromglc + 
                    lc_lb_gfsad) == 3
    rm(lc_lb_esaglc, lc_lb_fromglc, lc_lb_gfsad); gc()
  } else {
    message('Not use GFSAD for other types.')
    lc_lb_fromglc <- fromglc_new == lc$id
    lc_lb_esaglc <- resample(esaglc_new == lc$id, 
                             lc_lb_fromglc,
                             method = 'near')
    # gfsad_mask <- resample(gfsad_new == 0,
    #                        lc_lb_fromglc, 
    #                        method = 'near')
    lc_lb_emb <- (lc_lb_esaglc + 
                    lc_lb_fromglc) == 2
    message('Skip the cropland and water area from gfsad.')
    lc_lb_emb <- lc_lb_emb * gfsad_mask
    rm(lc_lb_esaglc, lc_lb_fromglc); gc()
  }

@@

For Tanzania case, there is no Tundra and Snow/Ice type because Tanzania is a tropical country.

@@

# save out
dst_path <- 'data/intermid'
fn <- file.path(dst_path, 'lc_labels.tif')
writeRaster(labels, fn, 
              wopt = list(datatype = 'INT1U',
                          gdal=c("COMPRESS=LZW")))
fn <- file.path(dst_path, 'labels_extent.tif')
writeRaster(labels_ext, fn, 
              wopt = list(datatype = 'INT1U',
                          gdal=c("COMPRESS=LZW")))
```

#### Select out sub-grids
Because of the under representation of classes: Bareland, Wetland, and Built-up, so gather them together as others class to ensure the good classification result. In general, we selected grids under three groups:
- Group 1: Relatively pure class grids (400 * 6 = 2400).
- Group 2: Highlighted areas such as cities and rivers (random 300)
- Group 3: Random grids across the consensus label areas (random 300)

##### Group 1
```{r g1}
library(ggplot2)
library(tidyr)
library(parallel)
# Define the width of chips to cut
# In this case, 256 * 256 chip.
# The chip size is the trade-off between computing cost, 
# purity of ensemble labels, and the keep of geographic context.
width_chips <- 16
lc_labels <- rast('data/intermid/lc_labels.tif')
tiles <- st_read('data/geoms/tiles_nicfi.geojson',
                 quiet = T) %>% dplyr::select(tile)
ecozones <- st_read('data/geoms/ecozones_simple.geojson',
                    quiet = T)

tiles <- do.call(rbind, mclapply(1:nrow(tiles), function(n){
  tile <- tiles %>% slice(n)
  tiles_grids <- st_make_grid(tile, n = c(width_chips, width_chips)) %>% 
    st_sf() %>% arrange() %>% 
    mutate(tile = tile$tile,
           index = c(1:(width_chips^2)))
  do.call(rbind, lapply(1:nrow(tiles_grids), function(m){
    tile <- tiles_grids %>% slice(m)
    ids <- st_intersects(tile, ecozones) %>% unlist()
    tile %>% mutate(ecozone = ids[1])
  }))
}, mc.cores = 12))
tiles <- tiles %>% filter(!is.na(ecozone))
st_write(tiles, 'data/intermid/tiles_grids_ecozones_256.geojson')
tiles <- tiles %>% vect()

@@

class_nm <- c('Cropland', 'Forest', 'Grassland', 'Shrubland', 
              'Wetland', 'Water', 'Urban/Built up', 'Bareland')
tiles_grids_pure <- do.call(rbind, 
                            lapply(class_nm, 
                                   function(nm){
    message(nm)
    # Ensure the enough amount of labels,
    # use different threshold to choose labels,
    # even though the labels might not be perfect.
    # Bareland, the most corrupted, so use 0.4
    # Shrubland and Cropland, use 0.7
    # Grassland, Forest, and water are probably the 
    # easiest classes.
    if (nm %in% c('Bareland', 'Wetland', 'Urban/Built up')) {
      tile_select <- tiles_labels %>% 
        dplyr::select(tile, index, all_of(nm)) %>% 
        filter(!is.na(!!rlang::sym(nm))) %>% 
        filter((!!rlang::sym(nm)) >= 0.6)
    } else {
      tile_select <- tiles_labels %>% 
          dplyr::select(tile, index, all_of(nm)) %>% 
          filter(!is.na(!!rlang::sym(nm))) %>% 
          filter((!!rlang::sym(nm)) >= 0.9)
    }
    data.frame(name = nm, 
               tile = tile_select$tile, 
               index = tile_select$index)
}))

grids_pure_count <- tiles_grids_pure %>% 
  group_by(name) %>% summarize(n = n()) %>% 
  arrange(n)
write.csv(grids_pure_count, 'data/intermid/pure_grids_count.csv',
          row.names = F)
```

```{r, eval=T, echo=F, message=F, warning=F}
grids_pure_count <- read.csv(here('data/intermid/pure_grids_count.csv'),
                             stringsAsFactors = F)

# Based on the table, the first three minor classes 
grids_pure_count %>%
  kbl(caption = "Pure label grids summary") %>%
  kable_classic(full_width = F, html_font = "Cambria")
```

```{r g1_sample}
set.seed(123)
grids_common_class <- tiles_grids_pure %>% 
  mutate(name = ifelse(name %in% c('Bareland', 'Wetland', 
                                   'Urban/Built up'), 'Other', 
                       name)) %>% 
  group_by(name) %>% 
  sample_n(500) %>% 
  ungroup()
```

##### Group 2
```{r g2}
# choose cities based on the consensus built-up
nm <- 'Urban/Built up'
cities <- tiles_labels %>% 
        dplyr::select(tile, index, all_of(nm)) %>% 
        filter(!is.na(!!rlang::sym(nm))) %>% 
        filter((!!rlang::sym(nm)) >= 0.2) %>%
  anti_join(grids_common_class, by = c('tile', 'index'))

# choose tiles overlapped with rivers
tiles <- st_read('data/intermid/tiles_grids_ecozones_256.geojson')
rivers <- st_read('data/geoms/gis_osm_waterways_free_1.shp') %>% 
  filter(fclass == 'river')
set.seed(160)
tiles_rivers <- tiles %>% 
  slice(unique(unlist(st_intersects(rivers, tiles)))) %>% 
  anti_join(grids_common_class, by = c('tile', 'index')) %>% 
  anti_join(cities, by = c('tile', 'index')) %>% 
  sample_n(1000 - nrow(cities))
cities_rivers <- rbind(cities %>% dplyr::select(tile, index),
                       tiles_rivers %>% dplyr::select(tile, index) %>% 
                         st_drop_geometry()) %>%
  unique() %>% mutate(name = 'Special')
```

##### Group 3
```{r g3}
g1_g2 <- rbind(grids_common_class, cities_rivers)
tiles <- tiles %>% anti_join(g1_g2, by = c('tile', 'index'))

# Only keep the grids with labels
target_vars <- setdiff(names(tiles_labels), 
                       c('ecozone', 'tile', 'index'))
tiles_full <- tiles_labels %>% 
  mutate(ratio = rowSums(.[target_vars], na.rm = TRUE)) %>% 
  filter(ratio > 0.8)
tiles <- tiles %>% right_join(tiles_full, by = c('tile', 'index'))

set.seed(170)
random_sites <- tiles %>% sample_n(1000) %>% 
  dplyr::select(tile, index) %>% 
  st_drop_geometry() %>% 
  mutate(name = 'random')
```

##### All grids

```{r g123}
grids_selected <- rbind(g1_g2, random_sites)
rm(g1_g2, random_sites, cities, rivers, tiles_rivers, cities_rivers);gc()
```

@@

```{r split_pure}
set.seed(100)
grids_train <- grids_selected %>% 
  group_by(name) %>% sample_frac(0.7) %>% 
  ungroup() %>% dplyr::select(-name) %>% 
  unique() %>% mutate(usage = 'train') %>%
  dplyr::select(tile, index, usage)

grids_valid <- grids_selected %>% 
  anti_join(grids_train, 
            by = c('tile', 'index')) %>% 
  group_by(name) %>% ungroup() %>% 
  dplyr::select(-name) %>% 
  unique() %>% mutate(usage = 'validate') %>%  
  dplyr::select(tile, index, usage)

# In order to set the batch size easily for DL,
# We re-arrange the sites among train and validate
# Train 3510, so 3510 / 30 could be without remainder.
# Validate 1470, so 1470 / 10 could be without remainder.
set.seed(678)
ids <- sample(1:nrow(grids_valid), 
              size = (nrow(grids_valid) - (3510 - nrow(grids_train))) %% 10)
grids_valid <- grids_valid %>% slice(-ids)

set.seed(568)
ids <- sample(1:nrow(grids_valid), size = 3510 - nrow(grids_train))
grids_valid$usage[ids] <- 'train'

@@

write.csv(catalog_train_val, 
          'data/intermid/catalog_guess_train_val.csv',
          row.names = F)
```

@@

```
id          name
1       Cropland
2       Forest
3       Grassland
4       Shrubland
5       Other (Wetland 5, Urban/Built up 8, Bareland 9)
6       Water
```

@@

# Make the script consistent
tiles <- st_read('data/intermid/tiles_grids_ecozones_256.geojson',
                 quiet = T)
lc_labels <- rast('data/intermid/lc_labels.tif')

plt_path <- '/Volumes/elephant/plt_nicfi'
plt_nms <- list.files(plt_path, full.names = T)
labels_path <- '/Volumes/elephant/labels_guess'

mclapply(unique(catalog_train_val$tile),
         function(tile_nm){
  tiles_each <- catalog_train_val %>% 
    filter(tile == tile_nm)
  lc_labels_each <- crop(lc_labels, tiles %>% 
                           filter(tile == tile_nm) %>% 
                           st_union() %>% 
                           st_buffer(0.01) %>% 
                           st_sf() %>% vect())
  nm <- grep(tile_nm, plt_nms, value = TRUE)[1]
  temp <- rast(nm)[[1]] %>% 
    aggregate(fact = 4096/width_chips)
  vals <- matrix(1:(width_chips^2), 
                 nrow = width_chips, 
                 ncol = width_chips) %>% t()
  vals <- vals[nrow(vals):1, ]
  values(temp) <- vals
  lapply(1:nrow(tiles_each), function(n){
    tile_each <- tiles_each %>% slice(n)
    temp_each <- temp
    index <- tile_each$index
    temp_each[temp_each != index] <- NA
    temp_each <- terra::trim(temp_each) %>% 
      disaggregate(fact = 4096/width_chips)
    label_each <- resample(lc_labels_each, temp_each, method = 'near')
    label_each[label_each %in% c(8, 9)] <- 5
    label_each[label_each %in% c(7, 10)] <- 0

@@

plts <- merge(plt_quads_rain_season, plt_quads_dry_season, by = 'tile')
rm(plt_quads_rain_season, plt_quads_dry_season)
## Labels
lc_labels <- list.files('/Volumes/elephant/labels_guess',
                        full.names = T, pattern = '.tif$') %>% 
    data.frame(dir_label = .) %>% 
    mutate(name = str_extract(dir_label, 
                              '[0-9]+-[0-9]+_[0-9]+')) %>% 
    tidyr::separate(name, c('tile', 'index'), '_')
imgs <- merge(lc_labels, plts, by = 'tile')
# rm(plts, lc_labels)

catalog_train_val <- read.csv('data/intermid/catalog_guess_train_val.csv',
                                  stringsAsFactors = F) %>% 
    merge(imgs, by = c('tile', 'index')) %>% 
    mutate(name = paste(tile, index, sep = '-')) %>% 
    dplyr::select(name, usage, dir_gs, dir_os, dir_label)
write.csv(catalog_train_val, 'data/intermid/catalog_DL_guess_train_val.csv',
          row.names = F)
```

@@

```{r}
pred_labeled <- catalog_train_val %>% 
  mutate(tile = str_extract(str_extract(dir_gs, 
                                        '[0-9]+-[0-9]+.tif'), 
                            '[0-9]+-[0-9]+')) %>% 
  dplyr::select(tile, dir_gs, dir_os) %>% unique() %>% 
  mutate(with_label = 1)

set.seed(320)
pred_unlabeled <- st_read('data/geoms/tiles_nicfi.geojson') %>% 
  filter(!tile %in% pred_labeled$tile) %>% 
  sample_n(200) %>% 
  merge(plts, by = 'tile') %>% 
  st_drop_geometry() %>% 
  dplyr::select(tile, dir_gs, dir_os) %>% 
  mutate(with_label = 0)

pred_catalog <- rbind(pred_labeled, pred_unlabeled)
rm(pred_labeled, pred_unlabeled)
write.csv(pred_catalog, 'data/intermid/catalog_DL_guess_predict.csv',
          row.names = F)
```
